---
title: "2021 FHB Pipeline Automatic Image Annotation vs. Manual Image Annotation - Large Scale Annotation; Plot"
author: "Julian Cooper"
date: "2023-09-27"
output: html_document
---

```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

# Overview
Correlations between pipeline automatic image annotation (AIA) plot average and manual image annotation (MIA) plot averages from the large scale annotation (LSA) data set

# Load Libraries
```{r,print = FALSE}

library(ggplot2) # For plotting data
library(tidyverse) # For manipulating/analyzing data
library(stringr) # For regex
library(GGally) # For ggcor
library(irr) # For calculating ICC
library(ggridges)    # for making ridge plot
library(viridis)     # color scaling in plot
library(rmarkdown) # For formatting rmarkdown
library(reshape2) # For reshaping correlation matrix

```

# Compare AIA Plot Averages to MIA LSA Plot Averages Based on ALL 5 RATERS
Filter for gradable spike images <= 0.50
```{r}

# Load AIA (automatic image annotation) data
aia <- read.csv("/Users/jcooper/Desktop/thesis_research/fhb_mineral/FHB_phenotyping_pipeline/automatic_image_annotation/2021model/data_processing/data/wheat2021model_automatic_image_annotation.csv")

# Calculate plot average FHB based on gradable spikes
aia_plot_avg <- aia %>% 
  filter(gradability_prob <= 0.50) %>%
  group_by(ID) %>% # group by ID for plot average
  summarise(aia_avg_fhb = mean(fhb_percentage, na.rm = TRUE)) %>% # Get plot average
  ungroup()

# Load MIA LSA (manual image large scale annotation) data
lsa <- read.csv("/Users/jcooper/Desktop/thesis_research/fhb_mineral/FHB_phenotyping_pipeline/manual_image_annotation/large_scale_annotations/data/2022/wheat2022_mia_large_scale_annotations.csv")

# Average plot average FHB based on graded spikes from all 5 rater
lsa_plot_avg <- lsa %>%
  group_by(ID) %>% # group by ID
  summarise(lsa_avg_fhb = mean(fhb_perc, na.rm = TRUE)) %>% # Get plot average
  ungroup() %>%
  separate(ID, into = c('Location', 'row_plot', 'PlotID', "Rating_Date"), sep = "_", remove = FALSE) # separate ID back into location, row_plot, PlotID, and Rating_Date

# Merge MFR and AIA 
aia_lsa <- merge(aia_plot_avg, lsa_plot_avg, by = "ID")

# Plot AIA vs LSA Plot Aggregate Disease Scores, color by location
plot_aia_lsa <- aia_lsa %>%
ggplot(aes(x = lsa_avg_fhb, y = aia_avg_fhb, color = Location)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "blue") +  # Add linear regression line
  geom_abline() +
  labs(x = "Large Scale Annotation FHB %", 
       y = "Automatic Image Analysis FHB %") +
  xlim(0,100) +
  ylim(0,100) +
  theme_classic()
print(plot_aia_lsa)
ggsave("/Users/jcooper/Desktop/thesis_research/fhb_mineral/FHB_phenotyping_pipeline/automatic_image_annotation/2021model/aia_mia_lsa_plot/results/2021model_scatterplot_AIA_LSA.png", plot = plot_aia_lsa, width = 6, height = 4)

# Correlations between AIA plot average FHB based on gradable spikes and LSA plot average FHB based on graded spikes from all 5 rater
aia_lsa %>%  
  summarise(cor(aia_avg_fhb, lsa_avg_fhb))

# Correlations between AIA and LSA plot averages, by location
aia_lsa %>% 
  group_by(Location) %>%
  summarise(cor(aia_avg_fhb, lsa_avg_fhb))

```
