---
title: "2021 FHB Pipeline Automatic Image Annotation vs. Manual Field Ratings; Plot"
author: "Julian Cooper"
date: "2023-09-27"
output: html_document
---

```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

# Overview
Correlations between pipeline automatic image annotation (AIA) plot average and manual field rating (MFR) plot averages

# Load Libraries
```{r,print = FALSE}

library(ggplot2) # For plotting data
library(tidyverse) # For manipulating/analyzing data
library(stringr) # For regex
library(GGally) # For ggcor
library(irr) # For calculating ICC
library(ggridges)    # for making ridge plot
library(viridis)     # color scaling in plot
library(rmarkdown) # For formatting rmarkdown
library(reshape2) # For reshaping correlation matrix
library(ggpubr) # For arranging plots

```

# Compare AIA Plot Averages to MFR Plot Averages Based on ALL 5 RATERS
Filter for gradable spike images <= 0.50
```{r}

# Load AIA (automatic image annotation) data
aia <- read.csv("/Users/jcooper/Desktop/thesis_research/fhb_mineral/FHB_phenotyping_pipeline/automatic_image_annotation/2021model/data_processing/data/wheat2021model_automatic_image_annotation.csv")

# Calculate plot average FHB based on gradable spikes
aia_plot_avg <- aia %>% 
  filter(gradability_prob <= 0.50) %>%
  group_by(ID) %>% # group by ID for plot average
  summarise(aia_avg_fhb = mean(fhb_percentage, na.rm = TRUE)) %>% # Get plot average
  ungroup()

# Load MFR (manual field rating) data
mfr <- read.csv("/Users/jcooper/Desktop/thesis_research/fhb_mineral/FHB_phenotyping_pipeline/manual_field_ratings/data/2022/wheat2022_manual_field_rating.csv")

# Calculate plot average FHB based on ALL 5 RATERS
mfr_all_rater_plot_avg <- mfr %>%
  pivot_longer(cols = c("A", "B", "C", "D", "E"), values_to = "fhb_perc", names_to = "Rater_ID") %>% # Pivot long by rater ID
  group_by(ID) %>% # group by ID for plot average
  summarise(mfr_avg_fhb = mean(fhb_perc, na.rm = TRUE))  %>% # Get plot average
  ungroup()

# Merge MFR and AIA 
aia_mfr <- merge(aia_plot_avg, mfr_all_rater_plot_avg, by = "ID")

# Extract date and location from ID
aia_mfr$Location <- sub("^(.{3}).*$", "\\1", aia_mfr$ID)
aia_mfr$Rating_Date <- sub(".*(.{10})$", "\\1", aia_mfr$ID)
aia_mfr$Location_Date <- paste(aia_mfr$Location, aia_mfr$Rating_Date)

# Plot AIA vs MFR Plot Aggregate Disease Scores
plot_aia_mfr <- aia_mfr %>%
ggplot(aes(x = mfr_avg_fhb, y = aia_avg_fhb)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "blue") +  # Add linear regression line
  geom_abline() +
  labs(x = "Manual Field Rating FHB %", 
       y = "Automatic Image Analysis FHB %") +
  xlim(0,100) +
  ylim(0,100) +
  theme_classic()
print(plot_aia_mfr)
ggsave("/Users/jcooper/Desktop/thesis_research/fhb_mineral/FHB_phenotyping_pipeline/automatic_image_annotation/2021model/aia_mfr_plot/results/2021model_scatterplot_AIA_MFR.png", plot = plot_aia_mfr, width = 6, height = 4)

# Plot AIA vs MFR Plot Aggregate Disease Scores for each Location/Date
plot_aia_mfr <- aia_mfr %>%
ggplot(aes(x = mfr_avg_fhb, y = aia_avg_fhb)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "blue") +  # Add linear regression line
  geom_abline() +
  labs(x = "Manual Field Rating FHB %", 
       y = "Automatic Image Analysis FHB %") +
  xlim(0,100) +
  ylim(0,100) +
  theme_classic() +
  facet_grid(~Location_Date)
print(plot_aia_mfr)
ggsave("/Users/jcooper/Desktop/thesis_research/fhb_mineral/FHB_phenotyping_pipeline/automatic_image_annotation/2021model/aia_mfr_plot/results/2021model_scatterplot_AIA_MFR_location_date.png", plot = plot_aia_mfr, width = 6, height = 4)

# Correlations for model to plot averages based on all 5 raters
aia_mfr %>%
  summarise(cor(aia_avg_fhb, mfr_avg_fhb))

# Correlations for model to plot averages based on all 5 raters for location/dates with corresponding manual image annotations
aia_mfr %>%
  filter(Rating_Date %in% c("2022-07-18", "2022-07-28")) %>%
  summarise(cor(aia_avg_fhb, mfr_avg_fhb))

# Correlations for model to plot averages based on all 5 raters for each location/date
aia_mfr %>%
  group_by(Location_Date) %>%
  summarise(cor(aia_avg_fhb, mfr_avg_fhb))
```

# Compare AIA Plot Averages to MFR Plot Aggregate Disease from INDIVIDUAL RATERS
Filter for gradable spike images <= 0.50
```{r}

# Load AIA (automatic image annotation) data
aia <- read.csv("/Users/jcooper/Desktop/thesis_research/fhb_mineral/FHB_phenotyping_pipeline/automatic_image_annotation/2021model/data_processing/data/wheat2021model_automatic_image_annotation.csv")

# Calculate plot average FHB based on gradable spikes
aia_plot_avg <- aia %>% 
  filter(gradability_prob <= 0.50) %>%
  group_by(ID) %>% # group by ID for plot average
  summarise(aia_avg_fhb = mean(fhb_percentage, na.rm = TRUE)) %>% # Get plot average
  ungroup()

# Load MFR (manual fiel rating) data
mfr <- read.csv("/Users/jcooper/Desktop/thesis_research/fhb_mineral/FHB_phenotyping_pipeline/manual_field_ratings/data/2022/wheat2022_manual_field_rating.csv")

# Merge MFR and AIA 
aia_mfr <- merge(aia_plot_avg, mfr, by = "ID")

# Extract date and location from ID
aia_mfr$Location <- sub("^(.{3}).*$", "\\1", aia_mfr$ID)
aia_mfr$Rating_Date <- sub(".*(.{10})$", "\\1", aia_mfr$ID)
aia_mfr$Location_Date <- paste(aia_mfr$Location, aia_mfr$Rating_Date)

# Correlations between pipeline and raters for each location
paged_table(aia_mfr %>%
  group_by(Location_Date) %>%
  na.omit() %>%
  summarise(aia_A = cor(aia_avg_fhb, A, method = "pearson"),
            aia_B = cor(aia_avg_fhb, B, method = "pearson"),
            aia_C = cor(aia_avg_fhb, C, method = "pearson"),
            aia_D= cor(aia_avg_fhb, D, method = "pearson"),
            aia_E = cor(aia_avg_fhb, E, method = "pearson")))

# Iterate through each AIA/Rater combo and plot AIA vs. MFR
for (i in c(13:17)) {
  plot_aia_mfr <- aia_mfr %>%
ggplot(aes(x = aia_mfr[,i], y = aia_avg_fhb)) +
geom_point() +
geom_smooth(method = "lm", se = FALSE, color = "blue") +  # Add linear regression line
geom_abline() +
labs(x = paste("Manual Field Rating FHB %", colnames(aia_mfr)[i]),
y = "Automatic Image Analysis FHB %") +
xlim(0,100) +
ylim(0,100) +
theme_classic() +
facet_grid(~Location_Date)
print(plot_aia_mfr)
ggsave(paste("/Users/jcooper/Desktop/thesis_research/fhb_mineral/FHB_phenotyping_pipeline/automatic_image_annotation/2021model/aia_mfr_plot/results/individual_raters/2021model_scatterplot_AIA_MFR_",colnames(aia_mfr)[i], ".png") , plot = plot_aia_mfr, width = 6, height = 4)

}


```
# Disease Progression of AIA Plot Averages Comapred to MFR Plot Averages Based on ALL 5 RATERS
Filter for gradable spike images <= 0.50
```{r}

# Load AIA (automatic image annotation) data
aia <- read.csv("/Users/jcooper/Desktop/thesis_research/fhb_mineral/FHB_phenotyping_pipeline/automatic_image_annotation/2021model/data_processing/data/wheat2021model_automatic_image_annotation.csv")

# Calculate plot average FHB based on gradable spikes
aia_plot_avg <- aia %>% 
  filter(gradability_prob <= 0.50) %>%
  group_by(ID) %>% # group by ID for plot average
  summarise(aia_avg_fhb = mean(fhb_percentage, na.rm = TRUE)) %>% # Get plot average
  ungroup()

# Load MFR (manual field rating) data
mfr <- read.csv("/Users/jcooper/Desktop/thesis_research/fhb_mineral/FHB_phenotyping_pipeline/manual_field_ratings/data/2022/wheat2022_manual_field_rating.csv")

# Calculate plot average FHB based on ALL 5 RATERS
mfr_all_rater_plot_avg <- mfr %>%
  pivot_longer(cols = c("A", "B", "C", "D", "E"), values_to = "fhb_perc", names_to = "Rater_ID") %>% # Pivot long by rater ID
  group_by(ID) %>% # group by ID for plot average
  summarise(mfr_avg_fhb = mean(fhb_perc, na.rm = TRUE))  %>% # Get plot average
  ungroup()

# Merge MFR and AIA 
aia_mfr <- merge(aia_plot_avg, mfr_all_rater_plot_avg, by = "ID")

# Extract date and location from ID
aia_mfr$Location <- sub("^(.{3}).*$", "\\1", aia_mfr$ID)
aia_mfr$Rating_Date <- sub(".*(.{10})$", "\\1", aia_mfr$ID)
aia_mfr$Plot <- sub(".*(76\\d{3}).*", "\\1", aia_mfr$ID)


# Filter for plots with disease progression data - StP
aia_mfr_StP <- aia_mfr %>%
  filter(Location == "StP")

# Add genotype
geno <- read.csv("/Users/jcooper/Desktop/thesis_research/fhb_mineral/FHB_phenotyping_pipeline/manual_field_ratings/data/raw_data/wheat2022_HTP_plot_geno_key.csv")
aia_mfr_StP <- merge(aia_mfr_StP, geno, by = "Plot")

# Creating a new data frame with correlations between mfr_avg_fhb and aia_avg_fhb for each Rating_Date
cor_df <- aia_mfr_StP %>%
  group_by(Rating_Date) %>%
  summarise(correlation = cor(mfr_avg_fhb, aia_avg_fhb))

# Heatmap of MFR disease progression
mfr_heat <- aia_mfr_StP %>%
  ggplot(aes(x = Rating_Date, y = Plot, fill = mfr_avg_fhb)) +
  geom_tile() +
  scale_fill_gradient(low = "blue", high = "red") +
  labs(x = "Rating Date",
       y = "",
       fill = "FHB %"
       ) +  # Empty y-axis label to remove y-axis labels
  facet_grid(Geno ~ ., scales = "free_y", space = "free_y", switch = "y") +  # Move facet titles to the left
  theme(strip.text.y.left = element_text(angle = 0),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        legend.position = "left") +
  scale_x_discrete(expand = c(0, 0))
print(mfr_heat)

ggsave("/Users/jcooper/Desktop/thesis_research/fhb_mineral/FHB_phenotyping_pipeline/automatic_image_annotation/2021model/aia_mfr_plot/results/mfr_heatmap.png", mfr_heat, width = 6, height = 5)

# Heatmap of AIA disease progression
aia_heat <- aia_mfr_StP %>%
  ggplot(aes(x = Rating_Date, y = Plot, fill = aia_avg_fhb)) +
  geom_tile() +
  scale_fill_gradient(low = "blue", high = "red") +
  labs(x = "Rating Date",
       y = "") +  # Empty y-axis label to remove y-axis labels
  facet_grid(Geno ~ ., scales = "free_y", space = "free_y", switch = "y") +  # Move facet titles to the left
  theme(strip.text.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        strip.background = element_blank(),
        legend.position = "none") +
  scale_x_discrete(expand = c(0, 0))
print(aia_heat)

ggsave("/Users/jcooper/Desktop/thesis_research/fhb_mineral/FHB_phenotyping_pipeline/automatic_image_annotation/2021model/aia_mfr_plot/results/aia_heatmap.png", aia_heat, width = 4.37, height = 5)


```
